Data- and Systems Paper : Finding Amsterdam

### 2016-05-27 milestone4/paper/paper.tex
anlaytics sort of maybe

As early as 2007, Schlobach et al. proposed approximation techniques based on logical assumptions of monotonicity. These techniques aimed to reduce the volume of data in order to improve on the running time \cite{schlobach2007anytime}

@@

%TODO done?
%Reasoning with large or complex terminology is computationally difficult and
%is one of the known bottlenecks in application of ontologies on the Semantic
%Web

\section{The dataset}
\textit{Laundromat}\footnote{\url{http://lodlaundromat.org/}} is a tool for cleaning data that provides well-formed datasets of different domains and volumes. Searching the LOD for sets linked to the OWL ontology, we found a set of ontologies that are expressive and complex enough to pose a challenge in the context of this assignment. Merging of the previous produced the \texttt{dataset} that we work on in this milestone. This \texttt{datset} was provided to us via StarDog.

The reasons why this dataset is challenging are the following:

\begin{enumerate}
	\item \textbf{Volume}: The given dataset is about 200 times larger than the ontology we created for the previous milestones.
	\item \textbf{Heterogeneity}: It is composed by several ontologies from different domains (foaf \footnote{Friend of a friend ontology \url{http://lov.okfn.org/dataset/lov/vocabs/foaf}}, wine, scientific-research, tourism).
	\item \textbf{Decentralization}: There is no central ontology that acts as an intermediate node for other ontologies to connect.
\end{enumerate}

Due to the above attributes, unforeseen challenges need to be tackled. These were not present in the well-designed dataset that we created previously.

\section{Data Analytics}
In order to get more detailed information regarding the \texttt{dataset} at hand, we performed an exploratory study of its most basic statistical characteristics. This is a X-step process that we describe now/here/something %TODO

Step 1: Query for finding the different ontologies

Step 2: Frequency of classes

To obtain frequency statistics we use the following query:

@@

138568 owl:Thing and 36273 owl:class

>>>>

As early as 2007, Schlobach et al. \cite{schlobach2007anytime} proposed approximation techniques based on logical assumptions of monotonicity. These techniques aimed to reduce the volume of data in order to improve on the running time, without loosing any or most of the logic entailment. Similar to us, they compared the selection of random axioms to two more sophisticated methods of selection. However, their procedure is based on frequency, either taking the most or least occurring entities, while we propose a degree approach, taking the best well-connected entities.

@@

%TODO done?

\section{The dataset}
\textit{Laundromat}\footnote{\url{http://lodlaundromat.org/}} is a tool for cleaning data that provides well-formed datasets of different domains and volumes. Searching the LOD for sets linked to the OWL ontology, one can find a set of ontologies that are expressive and complex enough to pose a challenge in the context of this assignment. Merging of some of the previous produced the \texttt{dataset} that we work on in this milestone. This \texttt{datset} was provided to us via the StarDog server.

At a first glance, the \texttt{dataset} is particularly challenging because of many varied reasons, mainly:

\begin{enumerate}
	\item \textbf{Volume}: The given dataset is about 200 times larger than the ontology we created for the previous milestones.
	\item \textbf{Heterogeneity}: It is composed by several ontologies from different domains (e.g the foaf \footnote{Friend of a friend ontology \url{http://lov.okfn.org/dataset/lov/vocabs/foaf}}, a wine ontology, a scientific-research ontology and a tourism ontology). Furthermore it comprises several different languages.
	\item \textbf{Decentralization}: There is no central ontology that acts as an intermediate node for other ontologies to connect.
\end{enumerate}

Due to the above attributes, unforeseen challenges need to be tackled. These were not present in the well-designed dataset that we created previously, therefore the reasoning methods we used before are not sufficient and need to be extended.

\section{Data Analytics}
Before reasoning or extracting any valuable information about the \texttt{dataset}, we first need to delve deeper into it to understand its content. In order to get more details, we performed an exploratory study of its most basic statistical characteristics. We did this in a 3-step process the we describe in the following subsections.

\subsection{Step 1: Query for finding the different ontologies}
First we needed to characterize the content of the \texttt{dataset} by the domains it refers to. To do so we query the ontologies that are present

\begin{lstlisting}[captionpos=b, caption=SPARQL query for getting different Ontologies in the \texttt{dataset}, label=lst:sparql, basicstyle=\ttfamily\small,frame=bt]
SELECT DISTINCT ?entity ?elabel ?type ?tlabel 
WHERE { 
	?entity a ?type .
	?type rdfs:label 'Ontology' .
	OPTIONAL { ?entity rdfs:label ?elabel } . 
	OPTIONAL { ?type rdfs:label ?tlabel } 
}
\end{lstlisting}

\subsection{Step 2: Frequency of classes}
To obtain frequency statistics we use the following query:

@@

The most frequent is owl:Thing with 138568 occurrences. Interestingly the second most frequent is not the owl:Class, with 36273 occurrences, but a class related to provenance as opmo:WasControlledBy with 38165 occurrences.
